---
title: "Homework 6"
author: "Nicholas Lai"
date: "November 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(astsa)
set.seed(123)
```
##Question 1

###(a)

```{r}
arma.spec(ar =c(1,-.9))
```

There is a dominant frequency between 0.1 and 0.2.

```{r}
sim_1 <- arima.sim(model = list(ar = c(1,-.9)), n = 50)
plot(sim_1)
```

There is periodic behavior of period between 5 and 10, as would be predicted by the dominant frequency in the spectral density.

###(b)

```{r}
arma.spec(ar=c(0,-0.3,0,-0.9))
```

There are two dominant frequencies, around 0.13 and 0.36.

```{r}
sim_2 <- arima.sim(model = list(ar = c(0,-0.3,0,-0.9)), n=50)
plot(sim_2)
```

We expect periodic behavior for period ~ 8 and 3, which is shown in this plot.

###(c)

```{r}
arma.spec(ar=c(0,0,0,0,0.9))
```

There are dominant frequencies around 0.2 and 0.4, seemlingly multiples of 0.2.

```{r}
sim_3 <- arima.sim(model = list(ar = c(0,0,0,0,0.9)), n = 50)
plot(sim_3)
```

From the spectral density we should expect periodic behavior around period 5, 2.5. This is clearly borne out by the plot.

##Question 2

###(b)

```{r}
arma.spec(ar = c(0,0,0.9))
```

The spectral density suggests that Xt will oscillate, around a period of 1/0.33, or 3.

###(c)
```{r}
sim <- arima.sim(model = list(ar = c(0,0,0.9)), n = 50)
plot(sim)

```

This is consistent with the spectral density above.

###(f)

```{r}
plot(filter(sim, filter = c(1/3,1/3,1/3), method = 'recursive'))
```

The model shows the same period as above, which is consistent with what we should expect from a linear filter.